Automated Verification for Real-Time Systems using Implicit Clocks and an Extended Antimirov Algorithm

### 2022-04-06 paper/tv.tex
m

We here present the forward verification rules, 
which are used to systematically accumulate the effects based on 
the syntax of each statement.
\code{\mathcal{P}} denotes the program being checked. 
With pre/post conditions declared for each method in \code{\mathcal{P}}, 
we can apply modular verification to a method's body using Hoare-style 
triples \code{\ \vdash \s{ \effect, \Pi, \env } \ e\ \s{\effect^\prime, \Pi^\prime, \env^\prime}}
where 
\code{\effect} is the current effects, 
\code{\Pi} is the current guard, 
\code{\env} is the current valuation of the variables;
and \s{\effect^\prime, \Pi^\prime, \env^\prime}  
is the resulting state by executing \code{e}. 
The program state is initialized using \s{\epsilon, True, []}.

@@

     \ |\ mn({v^*}) \ }

  |\ {if}\ b\ {then}\ e_1\ {else}\ e_2 \
 }  \\

 &   & & \code{
  |\ \textbf{event} [\anyevent{A}(v, \alpha^*)] 
  | \ \textbf{delay} [d] \ 
  | \ e_1\ \textbf{timeout} [d] \  e_2 \ 
  | \ \textbf{deadline}  [d] \ e \  
\end{comment}

\begin{gather*}
  \code{\frac{
  }
{  \vdash \{  \effect,\Pi, \env \}\  v \ 
\{ \effect, \Pi, \env \}} \  
 {[FV\text{-}Value]}} 
 \qquad
 \code{\frac{
  }
{  \vdash \{  \effect, \Pi, \env \}\  [b]\ e \ 
\{ \effect, \Pi{\wedge} b, \env  \}} \  
 {[FV\text{-}Guard]}} 
\end{gather*}

\begin{gather*}
  \code{\frac{ 
 \vdash \{ \effect, \Pi, \env\}\ e_1 \ \{  \effect^\prime, \Pi^\prime, \env^\prime\} 
\qquad 
\vdash \{ \effect^\prime, \Pi^\prime, \env^\prime\}\ e_2 \ \{  \effect^{\prime\prime}, \Pi^{\prime\prime}, \env^{\prime\prime} \} 
}{
  \vdash \{  \effect, \Pi, \env \}\  
  e_1;e_2 \ \{ \effect^{\prime\prime}, \Pi^{\prime\prime}, 
  \env^{\prime\prime}\}} \ 
   {[FV\text{-}Seq]}}  
\end{gather*}

\begin{gather*}
  \code{\frac{
 \vdash \{ \effect, \Pi, \env \}\ e_1 \ \{  \effect^\prime, \Pi^\prime, \env^\prime\} 
\qquad 
\vdash \{  \effect, \Pi, \env \}\ e_2 \ \{  \effect^{\prime\prime}, \Pi^{\prime\prime}, \env^{\prime\prime} \} 
    }{
\vdash \{   \effect, \Pi, \env \}\  e_1 || e_2 \ \{ \effect^\prime || \effect^{\prime\prime}, 
\Pi^{\prime} {\wedge}\Pi^{\prime\prime}, 
\}} \   
{[FV\text{-}Par]}} 
\end{gather*}

\begin{gather*}
  \code{\frac{\effect^\prime {=} \effect\cdot \anyevent{a}}{\vdash \{  \effect \}\  \textbf{event} [\anyevent{a}] \ \{ \effect^\prime \}} \   {[FV\text{-}Event]}} 
\end{gather*}

\begin{gather*}
\code{\frac{ \vdash \{  v \wedge \effect  \} \ e_1 \ \{  \effect^{\prime}\} \qquad \vdash \{ \neg v \wedge \effect \} \ e_2 \ \{  \effect^{\prime\prime}\}  }
{\vdash \{  \effect\}\  \textbf{if}\ v\ \textbf{then}\ e_1\ \textbf{else}\ e_2 \ \{  \effect^{\prime} \vee\effect^{\prime\prime} \} } \  {[FV\text{-}If\text{-}Else]}} 
\end{gather*}

\begin{gather*}
\code{\frac{
\begin{matrix}
\text{\code{  \tau \ mn  \ {( \tau\ x)^*}\ \{\textbf{requires}\ \effect_{pre}\ \textbf{ensures}\  \effect_{post}\} \ \{ e \} \in \mathcal{P}\ \ }}
\\
\text{\code{\vdash rev( \effect) \CONTAIN rev([y^*/x^*] \effect_{pre})   \qquad   \effect^\prime =  \effect \cdot  [y^*/x^*]\effect_{post}}}
\end{matrix}}
{\vdash \{  \effect \} \  mn({y^*})  \ \{  \effect^\prime\}} \ {[FV\text{-}Call]}}
\end{gather*}

\begin{gather*}
\code{\frac{  \vdash \{  \empt \}  \ e \  \{ \effect \} \qquad\qquad  \vdash\effect \CONTAIN   \effect_{post}}{  \vdash \tau \ mn  \ {( \tau\ x)^*}\ \{\textbf{requires}\ \effect_{pre}\ \textbf{ensures}\  \effect_{post}\} \ \{ e \}} \ {[FV\text{-}Meth]}}  
\end{gather*}

>>>>

We here present the forward verification rules, 
which are used to systematically accumulate the effects based on 
the syntax of each statement.
\code{\mathcal{P}} denotes the program being checked. 
With pre/post conditions declared for each method in \code{\mathcal{P}}, 
we can apply modular verification to a method's body using Hoare-style 
triples \code{\ \vdash \s{ \effect, \Pi } \ e\ 
\s{\effect^\prime, \Pi^\prime}}
where 
\code{\effect} is the current effects, 
\code{\Pi} is the current guard, 
%\code{\env} is the current valuation of the variables;
and \s{\effect^\prime, \Pi^\prime}  
is the resulting state by executing \code{e}. 
The program state is initialized using \s{\epsilon, True}.

@@

\begin{gather*}
  \code{\frac{
  }
{  \vdash \{  \effect,\Pi \}\  v \ 
\{ \effect, \Pi \}} \  
 {[FV\text{-}Value]}} 
 \qquad
 \code{\frac{
  }
{  \vdash \{  \effect, \Pi \}\  [b]\ e \ 
\{ \effect, \Pi{\wedge} b  \}} \  
 {[FV\text{-}Guard]}} 
\end{gather*}

\begin{gather*}
  \code{\frac{ 
 \vdash \{ \effect, \Pi\}\ e_1 \ \{  \effect^\prime, \Pi^\prime\} 
\qquad 
\vdash \{ \effect^\prime, \Pi^\prime\}\ e_2 \ \{  \effect^{\prime\prime}, \Pi^{\prime\prime} \} 
}{
  \vdash \{  \effect, \Pi \}\  
  e_1;e_2 \ \{ \effect^{\prime\prime}, \Pi^{\prime\prime}
  \}} \ 
   {[FV\text{-}Seq]}}  
\end{gather*}

\begin{gather*}
  \code{\frac{
 \vdash \{ \effect, \Pi \}\ e_1 \ \{  \effect^\prime, \Pi^\prime\} 
\qquad 
\vdash \{  \effect, \Pi \}\ e_2 \ \{  \effect^{\prime\prime}, \Pi^{\prime\prime} \} 
    }{
\vdash \{   \effect, \Pi \}\  e_1 || e_2 \ \{ \effect^\prime || \effect^{\prime\prime}, 
\Pi^{\prime} {\wedge}\Pi^{\prime\prime}
\}} \   
{[FV\text{-}Par]}} 
\end{gather*}

\begin{gather*}
  \code{\frac{ 
    \begin{matrix}
\vdash \{ \effect, \Pi {\wedge}  v   \} \ e_1 \ \{  \effect^{\prime} , \Pi^\prime \} 
\qquad \vdash \{ \effect, \Pi {\wedge} \neg v  \} \ e_2 \ 
\{  \effect^{\prime\prime}, \Pi^{\prime\prime} \}  
\\ \{  \effect_f, \Pi_f \} {=} \s{ \effect^{\prime} , \Pi^\prime} \cup \s{\effect^{\prime\prime}, \Pi^{\prime\prime}}
\end{matrix}}
  {\vdash \s{ \effect, \Pi}\  \textbf{if}\ v\ \textbf{then}\ e_1\ 
  \textbf{else}\ e_2 \ \{  \effect_f, \Pi_f \} } \  
  {[FV\text{-}If\text{-}Else]}} 
  \end{gather*}

\begin{gather*}
  \code{\frac{
    \effect^\prime {=} \effect\cdot \anyevent{A}(v, \Pi, \alpha^*) 
    }
  {\vdash \s{ \effect, \Pi} \  
  \textbf{event} [\anyevent{A}(v, \alpha^*)] 
   \ \{ \effect^\prime, True \}} \   {[FV\text{-}Event]}} 
   \qquad
  \code{\frac{
    \effect^\prime {=} \effect\cdot \epsilon\#t \quad (t\ is\ fresh) 
    }
  {\vdash \s{ \effect, \Pi} \  
  \textbf{delay} [d] 
   \ \{ \effect^\prime, True \}} \   {[FV\text{-}Delay]}} 
\end{gather*}

\begin{gather*}
  \code{\frac{
    \begin{matrix}
      \vdash \{ \epsilon, \Pi \}\ e_1 \ \{  \effect^\prime, \Pi^\prime\} 
      \qquad 
      \vdash \{  \epsilon, \Pi \}\ e_2 \ \{  \effect^{\prime\prime}, \Pi^{\prime\prime} \} 
      \\
      \s{\effect_f, \Pi_f} {=} \s{\effect {\cdot}(\effect^\prime\#t), \Pi^\prime {\wedge}t{\leq}d}
      \cup 
  \s{\effect {\cdot}(\effect^{\prime\prime}\#t), \Pi^{\prime\prime} {\wedge}t{>}d}   
  \quad (t\ is\ fresh)     
\end{matrix}
  }
  {\vdash \s{ \effect, \Pi} \  
  e_1\ \textbf{timeout} [d] \  e_2
   \ \s{\effect_f, \Pi_f} } \   {[FV\text{-}Timeout]}} 
\end{gather*}

\begin{gather*}
  \code{\frac{
    \begin{matrix}
      \vdash \{ \epsilon, \Pi \}\ e \ \{  \effect^\prime, \Pi^\prime\} 
      \qquad 
      \s{\effect_f, \Pi_f} {=} \s{\effect {\cdot}(\effect^\prime\#t), \Pi^\prime {\wedge}t{\leq}d}
  \quad (t\ is\ fresh)     
\end{matrix}
  }
  {\vdash \s{ \effect, \Pi} \  
  \textbf{deadline}  [d] \ e 
   \ \s{\effect_f, \Pi_f} } \   {[FV\text{-}Deadline]}} 
\end{gather*}

\begin{gather*}
\code{\frac{
\begin{matrix}
\text{\code{  mn  \ {x^*}\ \{\textbf{requires}\ \effect_{pre}\ \textbf{ensures}\  \effect_{post}\} \ \{ e \} \in \mathcal{P}\ \ }}
\\
\text{\code{\vdash  \effect \CONTAIN [v^*/x^*] \effect_{pre}   \qquad   \effect^\prime =  \effect \cdot  [y^*/x^*]\effect_{post}}}
\end{matrix}}
{\vdash \s{ \effect, \Pi} \ mn({v^*})  \ 
\{  \effect^\prime,  \Pi\}} \ {[FV\text{-}Call]}}
\end{gather*}

\begin{gather*}
\code{\frac{  \vdash \{  \empt, True \} 
 \ e \  \{ \effect, \Pi \} \qquad\qquad  \vdash\effect \CONTAIN  
  \effect_{post}}{  \vdash  \ mn  \ {x^*}\ \{\textbf{requires}\ \effect_{pre}\ \textbf{ensures}\  \effect_{post}\} \ \{ e \}} \ {[FV\text{-}Meth]}}  
\end{gather*}
